---
title: "Assignment MLogR FH"
author: "Ahmad Firdaus Mohamed"
date: "2023-12-19"
output: html_document
---

# Prepare Environment

## Load Library

```{r}
library(tidyverse)
library(gtsummary)
library(haven)
library(broom)
library(knitr)
library(tidyr)
library(ggplot2)
library(corrplot)
```

## Read data

```{r}
Data1<-read_sav("FraminghamTrue.sav")
```

# Data Wrangling

## Summary data

```{r}
summary(Data1)
```

```{r}
glimpse(Data1)
```

## Transform Data

```{r}
Data1 <- Data1 %>%
  mutate(across(where(is.labelled), as_factor)) %>%
  mutate(glucose = as.integer(glucose))

summary(Data1)
```

## Remove Missing Data

```{r}
#Data2 <- Data1 %>%
  #na.omit()
#summary(Data2)
```

```{r}
# Assuming your filename is "Data2.sav"
#file_path <- "C:/Users/drfir/OneDrive/Desktop/Data2.sav"

# Save the data frame to the specified file
#haven::write_sav(Data2, file_path)

```

## Summary in table

```{r}
Descriptive.tbl <- tbl_summary(by = TenYearCHD ,Data1)
Descriptive.tbl
```

```{r}
Descriptive.tbl2 <- tbl_summary(by = TenYearCHD, statistic = list(all_continuous() ~ "{mean}({sd})", all_categorical()~"{n} / {N} ({p}%)") ,Data1)
Descriptive.tbl2
```

## Explore Data

Histogram for numerical & barplot for categorical

Numerical : Age, CigPerday, totchol, SBP, DBP, BMI, HR, glucose

Categorical : Gender, Education, current smoker, BPmeds, PrevStroke, PrevHPT, DMstat, TenYearCHD

Age

```{r}
Data1 %>%
  ggplot(aes(age)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

CigPerday

```{r}
Data1 %>%
  ggplot(aes(cigsPerDay)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Total chol

```{r}
Data1 %>%
  ggplot(aes(totChol)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

SBP

```{r}
Data1 %>%
  ggplot(aes(sysBP)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

DBP

```{r}
Data1 %>%
  ggplot(aes(diaBP)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

BMI

```{r}
Data1 %>%
  ggplot(aes(BMI)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Heart Rate

```{r}
Data1 %>%
  ggplot(aes(heartRate)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Glucose

```{r}
Data1 %>%
  ggplot(aes(glucose)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Gender

```{r}
Data1 %>%
  ggplot(aes(gender)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Education

```{r}
Data1 %>%
  ggplot(aes(education)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Current Smoker

```{r}
Data1 %>%
  ggplot(aes(currentSmoker)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

BP Meds

```{r}
Data1 %>%
  ggplot(aes(BPMeds)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Prevalent Stroke

```{r}
Data1 %>%
  ggplot(aes(prevalentStroke)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Prevalent HPT

```{r}
Data1 %>%
  ggplot(aes(prevalentHyp)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

DM Status

```{r}
Data1 %>%
  ggplot(aes(diabetes)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Ten Year CHD

```{r}
Data1 %>%
  ggplot(aes(TenYearCHD)) + geom_bar()
```

## Checking Correlation

```{r}
#Data1 %>% 
 # select(-gender, -education, -currentSmoker, -BPMeds, -prevalentStroke, -diabetes, -TenYearCHD, -prevalentHyp) %>% 
  #ggpairs()
```

```{r}
Data2 <- 
  Data1 %>%
  select(where(is.numeric))
cor.Data2 <-
  cor(Data2, use = "complete.obs", method = "pearson")
head(round(cor.Data2,2))
corrplot(cor.Data2)
```

From the correlation plot, the SBP and DBP are highly correlated

```{r}
#cor.Data2 <-
#  cor(Data2, use = "complete.obs", method = "pearson")
#head(round(cor.Data2,2))
```

```{r}
#corrplot(cor.Data2)
```

## Causal Diagram

```{r}
myDAG <- dagitty( 'dag {
bb="0,0,1,1"
"Anti-Hypertensive Medication" [pos="0.602,0.881"]
"Smoking Status" [pos="0.743,0.775"]
"Total Cholesterol" [pos="0.302,0.122"]
Age [pos="0.541,0.072"]
BMI [pos="0.700,0.190"]
CHD [outcome,pos="0.817,0.482"]
Diabetes [exposure,pos="0.092,0.449"]
Gender [pos="0.185,0.754"]
Hypertension [pos="0.359,0.897"]
"Anti-Hypertensive Medication" -> CHD
"Anti-Hypertensive Medication" <-> Hypertension
"Smoking Status" -> CHD
"Smoking Status" -> Hypertension
"Total Cholesterol" -> CHD
"Total Cholesterol" -> Hypertension
Age -> CHD
Age -> Diabetes
Age -> Hypertension
BMI -> "Total Cholesterol"
BMI -> CHD
BMI -> Hypertension
BMI <-> Diabetes
Diabetes -> "Total Cholesterol"
Diabetes -> CHD
Diabetes -> Hypertension
Gender -> "Smoking Status"
Gender -> "Total Cholesterol"
Gender -> CHD
Gender -> Diabetes
Gender -> Hypertension
Hypertension -> CHD
}'
 )
plot( myDAG )
```

From the DAG, the exposure studied is Diabetes (diabetes) and the outcome of interest is developement of Coronoray Heart Disease (TenYearCHD).

We identified the confounders are Age (age), BMI (BMI) and gender (gender). Meanwhile the mediators are Total Cholesterol(totChol), Hypertension (prevalentHyp) and Anti-Hypertension Medication (BPMeds).

# **Estimation**

## Simple Logistic Regression

Based on literature review, these are the selected variable for study

Outcome : Ten Year Risk Developing CHD

Numerical : Age, total chol, BMI, sbp, dbp,glucose, CigsPerDay, heartRate

Categorical : HPT,DM, gender, current smoking status, BPMeds, education, prevalentStroke

```{r}
slr_CHD = glm(TenYearCHD ~ 1, data = Data1, family = binomial(link = "logit"))
summary(slr_CHD)
```

```{r}
slr_CHD0 = glm(TenYearCHD ~ 1, data = Data1, family = binomial(link = "logit")) 
add1(slr_CHD0, scope = ~ gender + age + education + currentSmoker + cigsPerDay + BPMeds + prevalentStroke + prevalentHyp + diabetes + totChol + sysBP + diaBP + BMI + heartRate + glucose,test = "LRT")
```

```{r}
slr.age <- glm(TenYearCHD ~ age, data = Data1, family = binomial(link = "logit"))
```

### Create Table Summary for SLogR

```{r}
tbl_uvregression(Data1, method = glm, y = TenYearCHD, method.args = list(family=binomial), exponentiate = TRUE) %>%
  add_n(location = "level") %>%
  bold_labels()%>%
  italicize_levels()
```

## Multiple Logistic Regression

### Model Without Interaction

```{r}
mlr.main <- glm(TenYearCHD ~ gender + age + cigsPerDay + prevalentHyp + diabetes + totChol + BMI + BPMeds , family = binomial(link="logit"), data=Data1)
summary(mlr.main)
tidy(mlr.main)
tidy(mlr.main, exponentiate = TRUE)
```

Table MLogR without interaction

```{r}
tbl_regression(mlr.main, exponentiate = TRUE) %>%
  add_n(location = "level") %>%
  bold_labels()%>%
  italicize_levels()
```

### Model with Interaction

Interaction chosen : total chol \* HPT

```{r}
mlr.main.ia <- glm(TenYearCHD ~ gender + age + cigsPerDay + prevalentHyp + diabetes + totChol + BMI + BPMeds + totChol:prevalentHyp , family = binomial(link="logit"), data=Data1)
summary(mlr.main.ia)
tidy(mlr.main.ia)
tidy(mlr.main.ia, exponentiate = TRUE)
```

Table MLogR with interaction

```{r}
tbl_regression(mlr.main.ia, exponentiate = TRUE) %>%
  add_n(location = "level") %>%
  bold_labels()%>%
  italicize_levels()
```

### Model Selection

Comparing using Anova

```{r}
anova(mlr.main,mlr.main.ia, test = 'Chisq')
```

From the ANOVA The p-value (Pr(\>F)) is greater than the typical significance level of 0.05. This suggests there is no difference between the two models.

Hence, we choose multivariable model without interaction (mlr.main) as our preliminary Final Model for Linear Regression due to model with the least variable is parsimonious model.

# Prediction

We will predict the :\
1. estimated log odds 2. probabilities 3. residuals 4. hat values 5. Cooks distance 6. standardized residuals

As of now, our preliminary final model is mlr.main

```{r}
prelim.final.m <- mlr.main
```

### Predict the log odds

```{r}
tidy(prelim.final.m, conf.int = TRUE)
```

### Predicted Odds

```{r}
tidy(prelim.final.m, exponentiate = TRUE, conf.int = TRUE)
```

### Fitted values

Probability of CHD

```{r}
augment(prelim.final.m, 
        type.predict = 'response') ##the type predict = response will change it from log odds terus jadi probaility. sama macam formula fitted kat bawah
```

Probality using Another Method

```{r}
prob.chd <- fitted(prelim.final.m)
head(prob.chd)
```

### Predict & Proving Manually the log-odds, odds, OR and probability

```{r}
summary(prelim.final.m)
```

From here, our equation will be

Log Odds = -7.139283 + 0.497630(gendermale) + 0.071868 (age) + 0.019649(cigsPerDay)+ 0.610469(prevalentHyp) + 0.724899(diabetes) + 0.002504(totChol) + 0.015248(BMI) + 0.349057(BPMeds)

```{r}
head(model.matrix(prelim.final.m))
```

For the first observation

1.  gendermale = 1

2.  age = 39

3.  cigsPerDay = 0

4.  prevalentHyp = 0

5.  diabetes = 0

6.  totChol = 195

7.  BMI = 26.97

8.  BPMeds = 0

```{r}
l.o.chd.link <- predict(prelim.final.m, type = 'link')
l.o.chd.term <- predict(prelim.final.m, type = 'term')
l.o.chd.prob <- predict(prelim.final.m, type = 'response')

##Link Scale (type = 'link'): Gives predictions on the log-odds scale.

#Linear Predictor (type = 'term'): Provides the raw linear combination of variables without transformation.

#Response Scale (type = 'response'): Gives predicted probabilities on the scale of the outcome variable.
```

```{r}
head(l.o.chd.link)
```

From here, the R calculated Log Odds for first observation is -2.9391689

```{r}
head(l.o.chd.term)
```

```{r}
head(l.o.chd.prob)
```

So to prove Log Odds for first observation =-2.9391689, we will use the equation above\

```{r}
-7.139283 + (0.497630*1) + (0.071868*39) + (0.019649*0)+ (0.610469*0) + (0.724899*0) + (0.002504*195) + (0.015248*26.97) + (0.349057*0)
```

So the odds for the first observation is\

```{r}
exp(-2.939282)
```

So, the probability for this is\
prob = odds/1+odds

```{r}
0.0529037 / (1+0.0529037)
```

```{r}
augment(prelim.final.m, 
        type.predict = 'response', 
        type.residuals = 'pearson')
```

### Prediction for new data

First, we create new data set based on the summary of our model

```{r}
summary(model.matrix(prelim.final.m))
```

\

```{r}
newdata1 <- expand.grid(age = seq(from = 32, to = 70, by = 10),
                       totChol = seq(from = 113, to = 400, by = 50),
                       diabetes = c('no dm', 'dm'),
                       gender = c('male', 'female'),
                       cigsPerDay = seq(from = 0, to = 70),
                       prevalentHyp = c('hpt', 'no hpt'),
                       BMI = seq(from = 15.54, to = 56.80, by=10),
                       BPMeds = c('not on hypertensive medication', 'on hypertensive medication'),
                       TenYearCHD = c('no chd', 'chd'))
```

And now, we can make prediction

-   using `type = 'link'` to generate predicted values for logits exp (Odss)

-   using `type = response` to generate predicted values based on probabilites

```{r}
mlr.main.newdata1 <- glm(TenYearCHD ~ age + totChol + diabetes + gender + cigsPerDay + prevalentHyp + BMI + + BPMeds, family = binomial(link = "logit"), data = newdata1)
summary(mlr.main.newdata1)
```

Predict

```{r}
predict(prelim.final.m, newdata = newdata1, type.predict = 'link')
```

```{r}
augment(prelim.final.m, newdata = newdata1)
```

We retest with our log odds equation for first observation in new data

Log Odds = -7.139283 + 0.497630(gendermale) + 0.071868 (age) + 0.019649(cigsPerDay)+ 0.610469(prevalentHyp) + 0.724899(diabetes) + 0.002504(totChol) + 0.015248(BMI) + 0.349057(BPMeds)

```{r}
-7.139283 + (0.497630*1) + (0.071868*32) + (0.019649*0)+ (0.610469*1) + (0.724899*0) + (0.002504*113) + (0.015248*15.54) + (0.349057*0)
```

# **Model checking**

## Model Checking

### Check overall fitness:

1.  Accuracy
2.  Sensitivity
3.  SPecificity

### Linearity in logits

by using:

• area under the curve

• Hosmer-Lemeshow test

• modidied Hosmer-Lemeshow test

• Oseo Rojek test

#### Hosmer-Lemeshow Test

```{r}
# Assuming 'model' is your logistic regression model
library(ResourceSelection)
hoslem.test(Data1$TenYearCHD, fitted(prelim.final.m), g = 10)  # Adjust 'g' as needed

```

P value is \<0.001, hence model shows good fit

#### Area under the Curve (ROC)

```{r}
# Assuming 'model' is your logistic regression model
library(pROC)
roc_curve <- roc(TenYearCHD ~ predict(prelim.final.m, type = "response"), data = Data1)
auc(roc_curve)
plot(roc_curve)

```

#### Diagnostic Plot

```{r}
plot(prelim.final.m)
```

From the Pearson residual plots, we can identify there are influential outliers greater than +2

The studentized residual plots, we can also noted there are outliers greater than +2, the leverage are higher than +2 suggests presence of influential outliers

Hence, these outliers observations may affect considerably our logistic regression model.

##### Identify Influentials outliers

Need to identify influentials obs,

```{r}
# Assuming you have a linear regression model named 'your_model' and the data frame is named 'original_data'
# your_model <- lm(y ~ x1 + x2, data = original_data)

# Calculate influence measures
infl <- influence.measures(prelim.final.m)
infl.val <- data.frame(infl$infmat)

# Set a threshold for Cook's distance deviations
cutoff_value <- 4/3656

# Filter out influential observations
filtered_data <- Data1 |> filter(infl.val$cook.d <= cutoff_value)


# Fit the logistic regression model again using the filtered data
mlr.filtered.influential <- glm(TenYearCHD ~ gender + age + cigsPerDay + prevalentHyp + diabetes + totChol + BMI + BPMeds, family = binomial(link="logit"), data = filtered_data)

# Compare model summaries
summary(mlr.filtered.influential)
summary(prelim.final.m)
```

```{r}
tidy(prelim.final.m, conf.int = TRUE)
tidy(prelim.final.m, exponentiate = TRUE)
```

```{r}
tidy(mlr.filtered.influential, conf.int = TRUE)
tidy(mlr.filtered.influential, exponentiate = TRUE)
```

### Reperform Model fitness

#### Hosmer-Lemeshow Test

```{r}
# Assuming 'model' is your logistic regression model
library(ResourceSelection)
hoslem.test(filtered_data$TenYearCHD, fitted(mlr.filtered.influential), g = 10)  # Adjust 'g' as needed
```

P value \<0.001, the filtered influential model maintains its model fitness.

#### Area under the Curve (ROC)

```{r}
# Assuming 'model' is your logistic regression model
library(pROC)
roc_curve.remove.inf <- roc(TenYearCHD ~ predict(mlr.filtered.influential, type = "response"), data = filtered_data)
auc(roc_curve.remove.inf)
plot(roc_curve.remove.inf)
```

Area under Receiver of Operating Characteristic Curve (AUC-ROC):

After model remodelling AUC-ROC is 82.26%, the model has improved to good discriminating effect.

### Reperform Diagnostic plots

```{r}
plot(mlr.filtered.influential)
```

Most residuals appear to be falling within greater -/+ 2, lesser influentials observation.

#### Compare between model without influential and preliminary final model(with influentials)

To identify the difference between B coefficients between these 2 models

```{r}
tidy(prelim.final.m, exponentiate = TRUE)
tidy(mlr.filtered.influential, exponentiate = TRUE)
```

Prelim Final Model --\> B for age is 1.07, where as Filtered Model --\> B for age is 1.14 Due to the influentials, there are significant changes of B under diabetes status.

# Interpretation

Hence, the final model will be:

```{r}
Final.Model <- mlr.filtered.influential
tidy(Final.Model, exponentiate = TRUE)
```

```{r}
tbl_regression(Final.Model, exponentiate = TRUE) %>%
  add_n(location = "level") %>%
  bold_labels() %>%
  italicize_levels() %>%
  as_gt() %>%
  gt::tab_header(title = "Table 3. Multiple Logistic Regression (Final Model)")
```

## Model Equation

Ten Years risk of CHD = (0.0000175) + 2.36(gender) + 1.14(age) + 1.02(cigsperday) + 2.12(prevalentHyp + 0.196(diabetes) + 1.00(totchol) + 1.02(BMI))

### Interpretation:

Gender, Age, Number of cigarette smoked per day, hypertension status have significant association to CHD.

1.  Men has 2.36 time the odds compared to women to develop CHD (95% CI:1.75,3.19) when adjusted to age, cigperday, hypertension status \@ being male has higher chance (36%) of developing CHD compared to female when adjusted to age, cigperday, hypertension status.
2.  Every 1 year increase of age, has 1.14 time the odds of developing CHD (95% CI: 1.11,1.16) when adjusted to gender, cigperday, hypertension status.
3.  Every 1 stick of cigarette per day, has 1.02 time the odds of developing CHD (95% CI 1.01,1.03) when adjusted to gender, age, hypertension status
4.  Those who diagnosed with hypertension, has 2.12 time the odds of developing CHD (95% CI 1.60,2.80) when adjusted to gender, age, cigperday.
5.  Diabetes status, total cholesterol level, BMI and on Hypertensive medications or not, are not statistically significant association to developing CHD after adjusted to all other variables.
