---
title: "Assignment MLogR FH"
author: "Ahmad Firdaus Mohamed"
date: "2023-12-19"
output: html_document
---

# Prepare Environment

## Load Library

```{r}
library(tidyverse)
library(gtsummary)
library(haven)
library(broom)
library(knitr)
library(tidyr)
library(ggplot2)
library(corrplot)
```

## Read data

```{r}
Data1<-read_sav("FraminghamTrue.sav")
```

# Data Wrangling

## Summary data

```{r}
summary(Data1)
```

```{r}
glimpse(Data1)
```

## Transform Data

```{r}
Data1 <- Data1 %>%
  mutate(across(where(is.labelled), as_factor)) %>%
  mutate(glucose = as.integer(glucose))

summary(Data1)
```

## Remove Missing Data

```{r}
#Data2 <- Data1 %>%
  #na.omit()
#summary(Data2)
```

```{r}
# Assuming your filename is "Data2.sav"
#file_path <- "C:/Users/drfir/OneDrive/Desktop/Data2.sav"

# Save the data frame to the specified file
#haven::write_sav(Data2, file_path)

```

## Summary in table

```{r}
Descriptive.tbl <- tbl_summary(by = TenYearCHD ,Data1)
Descriptive.tbl
```

```{r}
Descriptive.tbl2 <- tbl_summary(by = TenYearCHD, statistic = list(all_continuous() ~ "{mean}({sd})", all_categorical()~"{n} / {N} ({p}%)") ,Data1)
Descriptive.tbl2
```

## Explore Data

Histogram for numerical & barplot for categorical

Numerical : Age, CigPerday, totchol, SBP, DBP, BMI, HR, glucose

Categorical : Gender, Education, current smoker, BPmeds, PrevStroke, PrevHPT, DMstat, TenYearCHD

Age

```{r}
Data1 %>%
  ggplot(aes(age)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

CigPerday

```{r}
Data1 %>%
  ggplot(aes(cigsPerDay)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Total chol

```{r}
Data1 %>%
  ggplot(aes(totChol)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

SBP

```{r}
Data1 %>%
  ggplot(aes(sysBP)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

DBP

```{r}
Data1 %>%
  ggplot(aes(diaBP)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

BMI

```{r}
Data1 %>%
  ggplot(aes(BMI)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Heart Rate

```{r}
Data1 %>%
  ggplot(aes(heartRate)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Glucose

```{r}
Data1 %>%
  ggplot(aes(glucose)) + geom_histogram() + facet_grid(. ~TenYearCHD)
```

Gender

```{r}
Data1 %>%
  ggplot(aes(gender)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Education

```{r}
Data1 %>%
  ggplot(aes(education)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Current Smoker

```{r}
Data1 %>%
  ggplot(aes(currentSmoker)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

BP Meds

```{r}
Data1 %>%
  ggplot(aes(BPMeds)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Prevalent Stroke

```{r}
Data1 %>%
  ggplot(aes(prevalentStroke)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Prevalent HPT

```{r}
Data1 %>%
  ggplot(aes(prevalentHyp)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

DM Status

```{r}
Data1 %>%
  ggplot(aes(diabetes)) + geom_bar() + facet_grid(. ~TenYearCHD)
```

Ten Year CHD

```{r}
Data1 %>%
  ggplot(aes(TenYearCHD)) + geom_bar()
```

## Checking Correlation

```{r}
#Data1 %>% 
 # select(-gender, -education, -currentSmoker, -BPMeds, -prevalentStroke, -diabetes, -TenYearCHD, -prevalentHyp) %>% 
  #ggpairs()
```

```{r}
Data2 <- 
  Data1 %>%
  select(where(is.numeric))
cor.Data2 <-
  cor(Data2, use = "complete.obs", method = "pearson")
head(round(cor.Data2,2))
corrplot(cor.Data2)
```

From the correlation plot, the SBP and DBP are highly correlated

```{r}
#cor.Data2 <-
#  cor(Data2, use = "complete.obs", method = "pearson")
#head(round(cor.Data2,2))
```

```{r}
#corrplot(cor.Data2)
```

## Causal Diagram

```{r}
myDAG <- dagitty( 'dag {
bb="0,0,1,1"
"Anti-Hypertensive Medication" [pos="0.602,0.881"]
"Smoking Status" [pos="0.743,0.775"]
"Total Cholesterol" [pos="0.302,0.122"]
Age [pos="0.541,0.072"]
BMI [pos="0.700,0.190"]
CHD [outcome,pos="0.817,0.482"]
Diabetes [exposure,pos="0.092,0.449"]
Gender [pos="0.185,0.754"]
Hypertension [pos="0.359,0.897"]
"Anti-Hypertensive Medication" -> CHD
"Anti-Hypertensive Medication" <-> Hypertension
"Smoking Status" -> CHD
"Smoking Status" -> Hypertension
"Total Cholesterol" -> CHD
"Total Cholesterol" -> Hypertension
Age -> CHD
Age -> Diabetes
Age -> Hypertension
BMI -> "Total Cholesterol"
BMI -> CHD
BMI -> Hypertension
BMI <-> Diabetes
Diabetes -> "Total Cholesterol"
Diabetes -> CHD
Diabetes -> Hypertension
Gender -> "Smoking Status"
Gender -> "Total Cholesterol"
Gender -> CHD
Gender -> Diabetes
Gender -> Hypertension
Hypertension -> CHD
}'
 )
plot( myDAG )
```

From the DAG, the exposure studied is Diabetes (diabetes) and the outcome of interest is developement of Coronoray Heart Disease (TenYearCHD).

We identified the confounders are Age (age), BMI (BMI) and gender (gender). Meanwhile the mediators are Total Cholesterol(totChol), Hypertension (prevalentHyp) and Anti-Hypertension Medication (BPMeds).

# **Estimation**

## Simple Logistic Regression

Based on literature review, these are the selected variable for study

Outcome : Ten Year Risk Developing CHD

Numerical : Age, total chol, BMI, sbp, dbp,glucose, CigsPerDay, heartRate

Categorical : HPT,DM, gender, current smoking status, BPMeds, education, prevalentStroke

```{r}
slr_CHD = glm(TenYearCHD ~ 1, data = Data1, family = binomial(link = "logit"))
summary(slr_CHD)
```

```{r}
slr_CHD0 = glm(TenYearCHD ~ 1, data = Data1, family = binomial(link = "logit")) 
add1(slr_CHD0, scope = ~ gender + age + education + currentSmoker + cigsPerDay + BPMeds + prevalentStroke + prevalentHyp + diabetes + totChol + sysBP + diaBP + BMI + heartRate + glucose,test = "LRT")
```

```{r}
slr.age <- glm(TenYearCHD ~ age, data = Data1, family = binomial(link = "logit"))
```

### Create Table Summary for SLogR

```{r}
tbl_uvregression(Data1, method = glm, y = TenYearCHD, method.args = list(family=binomial), exponentiate = TRUE) %>%
  add_n(location = "level") %>%
  bold_labels()%>%
  italicize_levels()
```

## Multiple Logistic Regression

### Model Without Interaction

```{r}
mlr.main <- glm(TenYearCHD ~ gender + age + cigsPerDay + prevalentHyp + diabetes + totChol + BMI + BPMeds , family = binomial(link="logit"), data=Data1)
summary(mlr.main)
tidy(mlr.main)
tidy(mlr.main, exponentiate = TRUE)
```

Table MLogR without interaction

```{r}
tbl_regression(mlr.main, exponentiate = TRUE) %>%
  add_n(location = "level") %>%
  bold_labels()%>%
  italicize_levels()
```

### Model with Interaction

Interaction chosen : total chol \* HPT

```{r}
mlr.main.ia <- glm(TenYearCHD ~ gender + age + cigsPerDay + prevalentHyp + diabetes + totChol + BMI + BPMeds + totChol:prevalentHyp , family = binomial(link="logit"), data=Data1)
summary(mlr.main.ia)
tidy(mlr.main.ia)
tidy(mlr.main.ia, exponentiate = TRUE)
```

Table MLogR with interaction

```{r}
tbl_regression(mlr.main.ia, exponentiate = TRUE) %>%
  add_n(location = "level") %>%
  bold_labels()%>%
  italicize_levels()
```

### Model Selection

Comparing using Anova

```{r}
anova(mlr.main,mlr.main.ia, test = 'Chisq')
```

From the ANOVA The p-value (Pr(\>F)) is greater than the typical significance level of 0.05. This suggests there is no difference between the two models.

Hence, we choose multivariable model without interaction (mlr.main) as our preliminary Final Model for Linear Regression due to model with the least variable is parsimonious model.

# Prediction

We will predict the :\
1. estimated log odds
2. probabilities
3. residuals
4. hat values
5. Cooks distance
6. standardized residuals

As of now, our preliminary final model is mlr.main

```{r}
prelim.final.m <- mlr.main
```

###  Predict the log odds

```{r}
tidy(prelim.final.m, conf.int = TRUE)
```

### Predicted Odds

```{r}
tidy(prelim.final.m, exponentiate = TRUE, conf.int = TRUE)
```

### Fitted values

Probability of CHD

```{r}
augment(prelim.final.m, 
        type.predict = 'response') ##the type predict = response will change it from log odds terus jadi probaility. sama macam formula fitted kat bawah
```

Probality using Another Method

```{r}
prob.chd <- fitted(prelim.final.m)
head(prob.chd)
```

###  Predict & Proving Manually the log-odds, odds, OR and probability

```{r}
summary(prelim.final.m)
```

From here, our equation will be

Log Odds = -7.139283 + 0.497630(gendermale) + 0.071868 (age) + 0.019649(cigsPerDay)+ 0.610469(prevalentHyp) + 0.724899(diabetes) + 0.002504(totChol) + 0.015248(BMI) + 0.349057(BPMeds)

```{r}
head(model.matrix(prelim.final.m))
```

For the first observation

1.  gendermale = 1

2.  age = 39

3.  cigsPerDay = 0

4.  prevalentHyp = 0

5.  diabetes = 0

6.  totChol = 195

7.  BMI = 26.97

8.  BPMeds = 0

```{r}
l.o.chd.link <- predict(prelim.final.m, type = 'link')
l.o.chd.term <- predict(prelim.final.m, type = 'term')
l.o.chd.prob <- predict(prelim.final.m, type = 'response')

##Link Scale (type = 'link'): Gives predictions on the log-odds scale.

#Linear Predictor (type = 'term'): Provides the raw linear combination of variables without transformation.

#Response Scale (type = 'response'): Gives predicted probabilities on the scale of the outcome variable.
```

```{r}
head(l.o.chd.link)
```

From here, the R calculated Log Odds for first observation is -2.9391689

```{r}
head(l.o.chd.term)
```

```{r}
head(l.o.chd.prob)
```

So to prove Log Odds for first observation =-2.9391689, we will use the equation above\

```{r}
-7.139283 + (0.497630*1) + (0.071868*39) + (0.019649*0)+ (0.610469*0) + (0.724899*0) + (0.002504*195) + (0.015248*26.97) + (0.349057*0)
```

So the odds for the first observation is\

```{r}
exp(-2.939282)
```

So, the probability for this is\
prob = odds/1+odds

```{r}
0.0529037 / (1+0.0529037)
```

```{r}
augment(prelim.final.m, 
        type.predict = 'response', 
        type.residuals = 'pearson')
```

### Prediction for new data

First, we create new data set based on the summary of our model

```{r}
summary(model.matrix(prelim.final.m))
```

\

```{r}
newdata1 <- expand.grid(age = seq(from = 32, to = 70, by = 10),
                       totChol = seq(from = 113, to = 400, by = 50),
                       diabetes = c('no dm', 'dm'),
                       gender = c('male', 'female'),
                       cigsPerDay = seq(from = 0, to = 70),
                       prevalentHyp = c('hpt', 'no hpt'),
                       BMI = seq(from = 15.54, to = 56.80, by=10),
                       BPMeds = c('not on hypertensive medication', 'on hypertensive medication'),
                       TenYearCHD = c('no chd', 'chd'))
```

And now, we can make prediction

-   using `type = 'link'` to generate predicted values for logits exp (Odss)

-   using `type = response` to generate predicted values based on probabilites

```{r}
mlr.main.newdata1 <- glm(TenYearCHD ~ age + totChol + diabetes + gender + cigsPerDay + prevalentHyp + BMI + + BPMeds, family = binomial(link = "logit"), data = newdata1)
summary(mlr.main.newdata1)
```

Predict

```{r}
predict(prelim.final.m, newdata = newdata1, type.predict = 'link')
```

```{r}
augment(prelim.final.m, newdata = newdata1)
```

We retest with our log odds equation for first observation in new data

Log Odds = -7.139283 + 0.497630(gendermale) + 0.071868 (age) + 0.019649(cigsPerDay)+ 0.610469(prevalentHyp) + 0.724899(diabetes) + 0.002504(totChol) + 0.015248(BMI) + 0.349057(BPMeds)

```{r}
-7.139283 + (0.497630*1) + (0.071868*32) + (0.019649*0)+ (0.610469*1) + (0.724899*0) + (0.002504*113) + (0.015248*15.54) + (0.349057*0)
```
